---
title: "Acquisition of Spectra"
subtitle: "`ooacquire` `r packageVersion('ooacquire')`"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Acquisition of Spectra}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

## Summary

This package supports direct acquisition of spectra only from Ocean
Optics spectrometers. It also allows some computations on RAW spectral
counts to be done with the spectrometer off-line. What distinguishes it
from other software is the implementation of special algorithms for
measurement protocols that enhance the signal to noise ratio by one
order of magnitude, allowing the measurement of ultraviolet-B radiation
in sunlight using array spectrometers.

-   Allows spectral data acquisition from within an R session in near
    real-time.
-   Allows conversion of raw-counts data acquired using SpectraSuite or
    OceanView on PCs and possibly using the open-source SeaBreeze based
    software running on the Raspberry Pi single-board microcomputer. It
    also supports computations based on files acquired autonomously with
    the Jaz spectrometer.
-   Irradiance, fluence, reflectance, transmittance, absorptance and
    absorbance can be derived from raw spectra.
-   Measurements of both continuous and pulsed light sources is
    possible.
-   Different measurement protocols, data averaging, integration time
    bracketing, automatic adjustment of integration time can be used and
    corrections for stray light and slit function can be applied.
-   Acquisition of time series of spectra is also possible in two
    different ways: at timed intervals, or as fast as possible using
    buffered transfer from the spectrometer. The minimum delay depends
    on the spectrometer but can be as short as a few microseconds.

Although not directly supported by this package, the code it includes
can serve as a basis for equivalent protocols and corrections of data
acquired with software and/or spectrometers from other suppliers.

## Introduction

This document describes only high level functions for interactive data
acquisition. We first show simple examples of their use, assuming that
R, the needed R packages, the Temurin 8 **OpenJDK** or Oracle's Java 8
**JDK** and Ocean Optics' free OmniDriver runtime are all installed and
working (see this package's README and the README of package
'rOmniDriver' for details).

In this vignette we provide a tutorial on how to carry out measurements.
Please, see the vignette *Algorithms* for the details on the
computations.

The package exports both high level user-friendly functions and the
lower level functions used to build them. What we describe here is the
use of the higher level functions.

## Building this vignette

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(cache = FALSE, root.dir = system.file("extdata", package = "ooacquire"))
```

```{r}
# change this to TRUE to run acquisition examples
# these examples require user interaction to complete!!
sr.online <- FALSE
```

**NOTE:** This vignette can be built even if a spectrometer is not
connected as the code chunks with code that "talks" with the instrument
are by default not evaluated. Set above `sr.online <- TRUE` and connect
a spectrometer for which a calibration is available if you intend to
rebuild this vignette with live output from a measurement session.

## Measurement of light spectra

An emission spectrum of light emmited by a source or the spectrum of
light received on a surface describes the energy or photon distribution
as a function of wavelength. A calibration describes the relationship
between the signal response of a sensor and a physical quantity. For
spectral measurements this involves a description of the response of the
spectrometer to light of each measurable wavelength and a calibration of
detector pixels or monochromator positions to true wavelengths

Thus to acquire spectral irradiance, a suitable instrument calibration
and correction method description should be also available and in
practice implemented in software. Advanced algorithms implemented in
'ooacquire' require a special characterization of the spectrometer. If
such characterization is not available, a simpler approach, as normally
used with most other software, is also available in 'ooacquire'. If no
calibration is available in the format used by this package, an attempt
is made to retrieve calibration data from the connected spectrometer. In
most cases the wavelength calibration can be retrieved, but the
multipliers needed to compute spectral irradiance are stored in memory
only if expressly saved for this type of measurements as a result of a
calibration. An irradiance calibration is not needed to measure
quantities like spectral reflectance and transmittance that are relative
to a reference. Protocols, characterization and calibration information
can all be stored once in the special format used by 'ooacquire' if
available. Subsequently, the acquisition of spectra does not differ much
and any differences are reflected in the user interface as long as a
calibration is available. What differs is the quality of the spectral
data that is acquired.

In the code chunks below we use calibrations included as part of the
package. As the serial number stored with the calibration is validated
against that retrieved from the instrument, the examples below will work
only with our own instruments, rather than with the ones you may intend
to use. In most cases you will need to either import calibration data as
provided by Ocean Optics/Ocean Insight or for more complex correction
algorithms, manually create R objects with calibration and
characterisation data obtained from other sources. Such setup needs to
be done only once, and updated when the spectrometer is recalibrated.

When sunlight or radiation emitted by lamps driven by direct or
alternating current are measured the integration time can be used as
reference to compute fluxes such as energy irradiance or photon
irradiance. In the case of many lamps driven by alternating current
accurate measurements require us to consider the line frequency (50 Hz
in Europe, 60 Hz in USA) because light output will usually vary
cyclically at twice these frequencies (100 Hz or 120 Hz) making it
necessary to use integration times that are multiples of the durantion
of one half cycle of line frequency (10 ms and 8.3333 ms, respectively)
or alternatively use long-enough integration times for several cycles to
be included in the measurement interval. The same considerations apply
to transmittance, reflectance, and absorptance measurements using
continuous radiation sources.

When measuring radiation from pulsed light sources such as xenon flashes
we express the measurements as spectral power or photons per event
(pulse or "flash") and the integration time must simply be long enough
to encompass a known number of *whole* pulses. We use the number of
pulses to compute fluence rate per pulse or event. The same
considerations apply to transmittance, reflectance and absorptance
measurements using pulsed radiation sources.

The functions described below by default acquire for each measurement
event one or more raw-counts spectra from the spectrometer. They compute
the corresponding spectrum or spectra expressed in a physical quantity,
and save the data, both raw and computed, immediately, creating one
`.Rda` R-data file per measurement event. The operator is presented with
an annotated plot, with annotations that the user can adjust before
continuing for example choosing energy vs. photon units. The plot
corresponding to the measurement is saved to a `.pdf` file, also before
a new measurement is started. This approach to saving the data ensures
that in case of a computer failure or program crash, at most data for a
single measurement event will be lost.

Spectra grouped into collections can be also saved and plotted at any
time during a measurement session. When saving a collection all spectra
are collected into a `.mspct` object, a summary of the measurements is
computed and an index to the spectra compiled. All these R objects are
saved into a single `.Rda` file.

A measurement event can consist of a single *"light"* spectrum or of a
time series of *"light"* spectra, sharing the same *"dark"* and
*"filter"* spectra used as reference for their conversion into spectral
irradiance. Each raw- or instrument-counts spectrum can consist in one
or more scans or readings of the detector array, taken using identical
or different settings in the spectrometer. Once the desired protocol is
selected, the measurements can be acquired with little intervention from
the user, and the settings are retained across successive measurement
events, unless expressly modified.

It is possible to select a different quantity to be plotted and saved to
files by passing an argument to `qty.out`:

-   `"raw"` skips computations of spectra in physical units and saves to
    file only raw-counts data with metadata. In this conversion into
    physical units must be done at a later time. The descriptor of the
    instrument if missing is retrieved from the instrument.
-   `"cps"` (default if no spectral irradiance calibration is available)
    converts raw-counts into linearized counts-per-second and saves to
    file both raw-counts data with metadata and counts-per second
    spectral data. The descriptor of the instrument if missing is
    retrieved from the instrument.
-   `"irrad"` (default if a spectral irradiance calibration is
    available) converts raw-counts into corrected spectral irradiance (a
    flux rate expressed per unit time) and saves to a file both
    raw-counts data with metadata and spectral irradiance data. If the
    descriptor and spectral irradiance calibration are not available in
    R but available in the memory of the instrument they are retrieved.
-   `"fluence"` same as "irrad"\`, but the quantity computed is spectral
    fluence (a flux expressed per measuring event).

The saved data objects include metadata as attributes:

| purpose                                                                   | included  | accessor           |
|---------------------------------------------------------------------------|-----------|--------------------|
| user comment                                                              | optional  | `comment()`        |
| what was measured                                                         | optional  | `what_measured()`  |
| where was measured                                                        | optional  | `where_measured()` |
| when was measured in UTC time                                             | automatic | `when_measured()`  |
| simply how was measured                                                   | automatic | `how_measured()`   |
| instrument descriptor including calibration, serial number, configuration | automatic |                    |
| instrument settings used                                                  | automatic |                    |

: Attributes of spectral objects.

These attributes, together with saving raw counts in all cases, allows
recalculation of physical quantites and provides a built-in trace of the
origin of the data.

## Spectral irradiance

Spectral energy irradiance ($W m^{-2} nm^{-1}$) and spectral photon
irradiance ($mol s^{-1} m^{m-2} nm^{-1}$\$) are fluxes expressed as
energy or quanta, respectively per unit area of a receiving surface.
They can be inter-converted and there is no difference in how they are
measured or a need for different calibration.

After loading the package as shown in the next section below and
connecting a spectrometer by USB, calling function
`acq_irrad_interactive()` opens a connection to the spectroometer and
starts an interactive spectral-data acquisition session. The function
takes several arguments, which have as defaults the most frequently used
values. The `correction.method` and `descriptor` parameters are used to
pass the calibration and protocol information specific to a given
spectrometer. For spectrometers "known" to package 'ooacquire' these
also have suitable defaults which are selected after quering from the
spectrometer its serial number. These parameters are not totally
redundant as for a given spectrometer multiple correction algorithms,
calibrations for different entrance optics, may be available and
selected by overriding the defaults. In all cases these arguments are
checked to match the serial number of the attached spectrometer, and
rejected in cases of mismatch.

If this test fails or no calibration information is available and the
user has not explicitly supplied the information as arguments to the
call, an attempt is made to retrieve a calibration from the spectrometer
EEPROM. Because of the way in which irradiance calibrations are stored
in Ocean Optics spectrometers, calibration data retireved from the
EEPROM can only be used if the user provides information about the
diffuser used as entrance optics, and assuming that the calibration is
valid, thus among other things, done using the same optical fibre as in
being used for the measurements. If a calibration cannot be retrieved
from the EEPROM or information about the diffuser is not supplied, this
last attempt also fails, the function falls back into a mode that
returns counts-per-second instead of spectral irradiance, with a
warning. A wavelength calibration remains as a requirement but it rarely
missing.

### Acquiring a single irradiance spectrum

#### A simple example

Assuming a calibration is available the minimal steps to measure
spectral irradiance are as follows.

0 If R and RStudio are not already installed, install their latest
versions. If not already installed, install the 64 bit Java 8 JDK (Java
development kit; the JRE runtime is not enough). Install the OmniDriver
runtime. Install package 'rJava' and test that it is working, by running
some of the examples in its documentation. Install from CRAN package
'ggspectra' that will install most of the necessary dependencies.
Finally install packages 'rOmniDriver' and 'ooacquire', preferably by
anabling the repository located at (<https://r.r4photobiology.info>)
(see the README files for instructions).

1 Start R and load the package. We also change the default reporting of
warnings by R so that they are printed without the delay.

```{r}
library(ooacquire)

# print warnings at the time they are triggered
options(warn = 1)
```

2 Connect an Ocean Optics spectrometer to a USB port. To avoid problems
use a short and high quality USB cable, if this is not possible use an
active USB extension cable. In all cases use a short and known good USB
cable when using 'ooacquire' or a given spectrometer for the first time
and later keep this short and good USB cable always at hand to debug any
possible USB connection problems if they happen.

3 Start an interactive measuring session, in this example accepting all
defaults except for the choice of a simplified user interface.

```{r, eval = sr.online}
acq_irrad_interactive(interface.mode = "simple")
```

If everything works as expected, the version of OmniDriver that has been
found will be reported. If one or more spectrometers are found, they
will be listed with their serial numbers for the user to choose one of
them. Once a connection and software are initialized data acquisition is
possible.

4.  Follow the prompts to acquire spectra.

Default settings allow flexibility as settings can be interactively set,
but this leads to a rather verbose interface. If one knows in advance
what settings will use, these can be set when starting the session by
passing arguments that become the new defaults.

The code above is run at the R console or by sourcing a simple script. I
normally use the RStudio IDE, but this is not required as long as
graphical display is enabled. The statements can be saved into an R
script (as a plain-text file) and then the file *sourced*. As scripts
can be also run from the operating system shell or command prompt, it is
possible, if desired to partly hide this step from users' view by
writing a batch file or shell script.

Of the formal parameters of function `acq_irrad_interactive()`,
`interface.mode` selects different sets of settings to be controlled
interactively during a session. The difference between the interface
modes is mostly in which of the different possible settings can be
modified by the user at run time (in other words, it alters which
options are visible or not in menus as well as some of the default
settings). The simplified interfaces make measurements simpler and
faster in specific situations, while the defaults for the session can
always be changed by passing arguments when calling the function. The
default `interface.mode = "auto"` retains the behaviour of earlier
versions, but in many cases `interface.mode = "simple"` will still give
access to all the parameters that need to be set at runtime for common
measurements. Modes `"manual"` and `"full"` make it possible to manually
set the integration time, while "series" makes it possible to acquire
time series of spectra as a single measurement event.

In all cases when values supplied by the user fall outside those
accepted by the connected instrument, these values will be adjusted to
the nearest valid value.

We can pass the objects containing the correction method definition and
the instrument descriptor that includes the calibration data as
arguments.

```{r, eval=sr.online}
acq_irrad_interactive(correction.method = ooacquire::MAYP11278_ylianttila.mthd,
                      descriptors = ooacquire::MAYP11278_descriptors)
```

In this case, this corresponds to a spectrometer whose data are
available in the package and recognized automatically. Thus is
equivalent to accepting all defaults as below.

```{r, eval=sr.online}
acq_irrad_interactive()
```

However, for this spectrometer additional correction methods are
available. Two calibrations with different entrance optics are
available, and can be selected by means of an argument passed to
parameter `entrance.optics`.

In the case of calibrations done by Ocean Insight or with their
software, and retrieved from the spectrometer EEPROM we need to pass an
argument to parameter `area` indicating the effective area of the
diffuser, or the one of one of the types known to 'ooacquire'.

The interface mode can be also selected only when call the function. The
default is `"auto"` for automatic integration-time tuning, `"manual"`
for manual setting of the integration time, `"series"` for the
acquisition of time series of spectra. Mode `"simple"` is a simplified
version of `"auto"`.

Although we discuss settings in relation to arguments to parameters, the
settings discussed below can be also set interactively by the user, at
least in one of the supported interface modes. The functions described
here, except when in interface mode `"manual"`, will set a suitable
integration time automatically.

#### Protocols for enhanced performance

Dynamic range is the difference between the smallest and largest values
that can be detected in a measurement. Noise is the component of the
measured value that is not a reflection of what we want to measure. An
array detector not only responds to light, but it is also to a small
extent excited also by thermal radiation, electrical fields and even
cosmic particles. These events are infrequent and so during a single
integration are likely to affect only individual pixels and different
pixels in different integrations. A further problem is stray light,
photons that imping on the "wrong" pixel because of internal reflections
in the optical bench of the spectrometer. Different protocols and
correction algorithms can mitigate the effect of these problems on the
quality of the spectral data.

If we use a single value for integration time (the equivalent of
exposure time in a photographic camera) the realized dynamic range is
the intrinsic dynamic range of the array detector and electronics of the
spectrometer. In this case we pass a single numeric value as argument to
`HDR.mult` to indicate acquisition using a single integration time.
Except in very unusual cases, we use 1, because `HDR.mult` is a
multiplicative factor applied to the optimal integration time.

```{r, eval=sr.online}
acq_irrad_interactive(HDR.mult = 1,
                      tot.time.range = c(0, Inf))
```

Averaging of multiple "scans" mitigates "random" noise because the
longer the total measurement time the most likely it is that all pixels
are similarly affected by the noise. The noise is still present, but it
affects more evenly the different pixels. The approach used in
`acq_irrad_interactive()` is for the user to supply an upper and lower
limit to the sum of integration times from multiple scans and the
software determines the number of scans that are needed at the time the
optimal integration time is assessed. Keeping the `tot.time.range`
constant can help keep the noise levels constant.

The argument passed to `tot.time.range`, a vector of one or two numbers,
indicates the allowed range for total time in seconds. If two identical
numbers or a single number are passed, this exact time is used. The
integration time for individual scans is adjusted automatically to
ensure good performance of the instrument and used to compute the number
of scans to average. The maximum and minimum integration times supported
by the spectrometer hardware set the effective limits for the
integration time. The smallest argument accepted for `tot.time.range` is
zero and the largest `Inf`. Setting `tot.time.range = c(0, Inf)` as in
the example above ensures that only one integration per measurement will
be used, and that the integration times used will be from the whole
range the spectrometer supports.

It would be possible to use in the example above a value larger than one
for `HDR.mult` even if a single integration time is used. This is
unusual, but if we indeed would like to force clipping of part of the
spectrum, this would be the way to achieve this automatically relative
to the optimum setting.

The dynamic range in the spectral data can be increased by splicing
spectra measured using increasing integration times (similar to
"exposure bracketing" as used in photography followed by merging to
ensure enough detail is captured in both shade and highlights). In most
cases for this to significantly improve the quality of the spectral data
we also need to average data from several successive integrations to
control the noise. The default in `acq_irrad_interactive()` are thus to
use the optimal integration time plus one ten times longer and to keep
the total measuring time, the sum of multiple integration events,
between 5 and 10 s.

```{r, eval=sr.online}
acq_irrad_interactive(HDR.mult = c(short = 1, long = 10), # the default
                      tot.time.range = c(5, 15), # the default
                      correction.method = ooacquire::MAYP11278_ylianttila.mthd,
                      descriptors = ooacquire::MAYP11278_descriptors)
```

The resulting spectra are merged choosing the *most suitable* of the
used integration times for each wavelength region. Usually using two
integration times with 1 and 10 as `HDR.mult` is a good compromise.
However, there is no built-in limitation in the package code making it
possible to use a vector of two, three, four or more values, which
combined with a very long total acquisition time of the order of
minutes, can yield very good control of noise for light sources with
stable output. Consequently, although the default setting
`HDR.mult = c(1, 10)` tends to work well, settings like
`HDR.mult = c(1, 5, 25)` may be useful in special circumstances.

We next set total acquisition time to a fixed length of 10 seconds. Once
this setting is active, each time the integration time is automatically
set, both the integration time and number of scans are adjusted so that
total acquisition time is exactly 10 seconds, for both short and long
integration times. Averaging multiple scans smooths out noise but
retains the spectral resolution. A constant total measurement time is
useful as it controls the influence of "noise" on the acquired data
consistently irrespective of the integration time used.

```{r, eval=sr.online}
acq_irrad_interactive(HDR.mult = c(1, 10),
                      tot.time.range = 10)
```

If long integration times are needed, the setting above may result in
the use of suboptimal integration times, as the desired integration time
may have be to be shortened until it becomes an exact fraction of the
total acquisition time. To avoid this, we can supply instead of a single
fixed value for the total acquisition time, a range of values.

```{r, eval=sr.online}
acq_irrad_interactive(HDR.mult = c(1, 10), 
                      tot.time.range = c(10, 20))
```

Sometimes we may want to avoid long integration times even at the cost
of not using the whole dynamic range of the detector. For example, when
following some fast kinetics, we would usually need the integration time
not to exceed a certain time, say 50 milliseconds, we would simply use
smaller values, such as `tot.time.range = c(0.03, 0.05)`.

One problem with automatic setting of integration time is that depending
on the stability of the light source we will need to set a different
value for the target relative to the longest integration time that would
result in no clipping during automatic tuning. Parameter `target.margin`
allows the user to set a *safety margin* different from the default of
0.1 (or 10%). The example below could help prevent clipping in cases
when irradiance changes so fast that it could be up to 25% higher during
actual measurement than during the preceding tuning of the integration
time.

```{r, eval=sr.online}
acq_irrad_interactive(target.margin = 0.25)
```

Although above we described the use of `HDR.mult` as a step in splicing
spectra, it is also possible at a later time also the discard spectra
acquired using specific multipliers and recalculate the irradiance. In
some cases with extremely variable light levels when acquiring time
series of spectra, it can be wise to use a rather wide range of values
in the vector passed to `HDR.mult`, which also accepts values smaller
than one, achieving a similar effect as `target.margin`
(`target.margin = 0.25` is equivalent to `HDR.mult = 0.75`).

The automatic tuning step can be also skipped. For example the value
used in the previous measurement event can be reused.

None of the approaches described above can control stray light caused by
reflections in the optical bench. Correction for stray light in the case
of some spectrometers makes it necessary to measure it using a short
pass filter. The function also supports this type of protocols with two
references, and uses it for two spectrometers to allow the measurement
of UV-B radiation in sunlight.

### Acquiring a time series of irradiance spectra

Measuring a time series of irradiance spectra using the approaches
described above can be difficult. Triggering acquisition of spectra at
regular intervals manually is difficult for long time steps and
impossible for very short ones. In most cases we also need to measure
many light spectra for each "dark" reference measurement. Interface mode
`"series"` introduced in version 0.3.0 of 'ooacquire' and enhanced in
laters versions targets this type of measurements.

For measuring a time series, we pass `interface.mode = "series"` in the
call to `acq_irrad_interactive()`. We can set the default parameters for
the series through arguments. The following example sets the default for
a series with measuring events triggered at the start of each minute,
with 60 s time step between successive `"light"` spectra and 60 spectra,
for a total time expanse of 1 h. We also set the initial delay to zero,
so that once we start the light measurement event by pressing enter, the
acquisition of the time series starts at the beginning of the next
minute.

```{r, eval=sr.online}
acq_irrad_interactive(interface.mode = "series",
                      seq.settings = list(start.boundary = "minute",
                                          initial.delay = 0, # seconds
                                          step.delay = 60, # seconds
                                          num.steps = 60))
```

More realistically, we would set some other defaults. The example below
has worked well in sunlight.

```{r, eval=sr.online}
acq_irrad_interactive(interface.mode = "series",
                      tot.time.range = c(5, 10),
                      target.margin = 0.1,
                      qty.out = "irrad",
                      HDR.mult = c(0.3, 1, 3, 10),
                      seq.settings = list(start.boundary = "minute",
                                          initial.delay = 0,
                                          step.delay = 60,
                                          num.steps = 60))
```

That the acquisition of spectra is continuously running in the
spectrometer is something to be aware of, because it means that this is
what determines the true time step between successively retrieved
spectra and that, say using a `step.delay` that is not a multiple of the
frequency at which the spectrometer is acquiring the spectra will result
only in missing the retrieval of some of the spectra measured by the
spectrometer, but not in a different effective time step! In other words
with short time steps we need to be aware that new spectra are acquired
at time steps that are a multiple of the **integration time** setting of
the spectrometer plus some small overhead that depends on the model of
the spectrometer. This is tricky as the length of the integration time
needs to be adjusted based on irradiance. The problem barely exists if
we average multiple spectra during a measurement, or the time step
includes some idle time between acquisitions.

The OmniDriver API contains special functions for acquisition of spectra
at high speed. These functions store the spectra in a computer memory
buffer for later retrieval by user code, minimizing the overhead, but
they do not allow control over the timing of successively acquired
spectra. The `"series"` mode uses them only for `step.delay = 0` and no
HDR integration time bracketing. In all other cases normal API functions
are used so as to allow the use of HDR bracketing or the choice of the
time step between successive acquisitions of spectra. Because, of how OO
array spectrometers work, some limitations exist. When using
`step.delay = 0` the time step is controlled by the length of the
measurement, or the integration time if spectra are not averaged for
each measurement.

In my test with a specific PC under Windows 10 together with a USB2000
spectrometer and without bracketing I was able to use a minimum
`step.delay = 0.052` with `tot.time.range = 0.050` with `HDR.mult = 1`.
From these tests I infer that the USB2000 spectrometer is constantly
measuring spectra with an overhead of $2\,ms$ (this delay is even
shorter in modern spectrometer models) and that the overhead incurred by
'ooacquire' in reading a spectrum is less than 50 ms.

In the memory buffered mode with a Maya 2000 Pro I was able to measure
using an integration time of $7.2\,ms$ a series with a median time step
of less than 7.23 ms or an overhead of less than $30\,\mu s$. However, I
observed some small variation in the length of succesive time step
across a series. Anyway time stamps are saved with a resolution of
$1\,\mu s$.

**The time overhead depends on the spectrometer, the PC and most likely
the operating system and possibly the version of OmniDriver, Java, R and 'ooacquire'.
Consequently before attempting measurements using `step.delay` values
that are not much longer than the largest value in the vector passed to
`tot.time.range`, some experimentation will be needed. The package
attempts to detect `step.delay` values that are too short to be
achieved and overrides them with `step.delay = 0`**.

When using HDR-integration-time bracketing, the spectrometer is not free running
or continuously acquiring spectra. Settings need to be changed, acquisition of
spectra restarted and the spectral data subsequently acquired only after a
freshly measured spectrum is available. This adds a lot of overhead compared to
retrieving spectra that are being measured by the spectrometer using
multitasking while the PC is busy with other tasks, such as retrieving the
previous spectrum. Furthermore, this also means that when using HDR bracketing,
the overhead depends on the `tot.time.range` and `HDR.mult` settings and on the
integration time. In contrast, when the settings are not modified between
successively acquired spectra the overhead is much less and less dependent on
the integration time and arguments passed to other parameters.

So with care about choosing matching integration times and `step.delay` values
without HDR bracketing one can achieve as little as 50 ms between acquired
spectra at timed steps. However, when using HDR bracketing, a rough guide is
that the minimum time step that can be achieved is about twice the sum of the
summed integration time values used. When `tot.time.range` is set as a range,
then the longest time should be considered as limiting.

The high speed acquisition mode retrieves as quickly as possible the spectra as
they are acquired, and the time step is simply the integration time plus the
spectrometer overhead. To enable this mode we use `step.delay = 0`. The time
stamp used for spectra in 'ooacquire' until version 0.3.0 has been the time when
spectra are retrieved from the spectrometer. This is close enough for most
individual measurements and series with time steps of several seconds, but when
dealing with fast time series this can be inadequate. If `step.delay = 0` is
used, the time for the first spectrum is set when the retrieval into the memory
buffer is started, but the times for the later spectra are set based on time
differences computed from time stamps retrieved from OmniDriver. The theoretical
time resolution is better than a microsecond for objects of R's class `POSIXct`,
and this class is used in package 'photobiology' to record acquisition times.

------------------------------------------------------------------------

**Measurement of series of spectra with single dark and filter
measurements makes it imperative that the functioning of the
spectrometer is stable. Consequently, when measuring series, specially
if they are longer than a few seconds, do make sure that the temperature
of the spectrometer has stabilized and that it remains stable throught
the acquisition of the series. For example, protect the spectrometer
from direct sunlight or other sources of energy that can increase its
temperature, while allowing ventilation to allow the dissipation of the
heat generated by the spectrometer itself. Of course, active temperature
control is the best approach. Passive temperature equalization can take
30 to 45 min if the difference between the storage temperature and the
temperature under which the spectrometer is used is large (e.g., more
than 10 C).**

**As the spectrometer electronics dissipate energy, even when used at
the same temperature at which it was stored, the spectrometer should be
allowed to warm up for 10 to 20 minutes before attempting to measure a
time series. Although allowing the spectrometer to warm up for several
minutes is always beneficial, when a single reference dark measurement
is used for a series of light measurements, it is crucial that the dark
reference is not invalidated by changes in the temperature of the
spectrometer.**

------------------------------------------------------------------------

### Spectral fluence

Function `acq_irrad_interactive()` has replaced function 
`acq_fluence_interactive`. Fluence spectra are returned by 
`acq_irrad_interactive()` when `qty.out = "fluence"`. Some of the defaults
for other parameters also change. Otherwise all the functionality described
above for irradiance measurements is avaialable except the `"series"` interface
mode.

If the duration of the exposure to a continuous light source is known, then
spectral irradiance multiplied by this time duration gives spectral fluence, and
can be obtained by measuring spectral irradiance as described in the previous
section. In this section we discuss the case when spectral fluence is measured
directly, using an integration time that encompasses the whole duration of the
irradiation event, especially when the duration of this event is unknown or
spectral irradiance varies through the irradiation event and we are interested
in the accumulated exposure or "dose".

Fluence, also called radiant exposure, is defined per exposure event
rather than per unit time. When measuring pulsed light sources, the
length of exposure is determined by the duration of the pulse, which is
usually not known with precision. In the discussion below we assume that
the quantity of interest is fluence per pulse. For this type of
measurements the integration time setting in the spectrometer only
affects the window within which pulses will be measured, consequently,
it is always set manually. The way of controlling the number of
photons reaching the detector during an integration is through the
number of pulses per integration and/or the fluence provided by each
pulse (e.g., by altering the distance between source and instrument 
entrance optics or the power settings of the source).

We start the session as described above for irradiance, but using
function `acq_fluence_interactive()`. Using defaults except for `qty.out`.

```{r, eval = sr.online}
acq_irrad_interactive(qty.out = "fluence")
```

The integration time can be set when calling the function to start a
session or leter interactively.

Another important parameter is the number of exposure events. This
determines how the resulting data are expressed. If we consider that
each individual pulse is an exposure event, and for example, we trigger
five flashes per spectrometer integration (or "scan"), we would set
`num.exposures = 5`.

By default the user is prompted to manually trigger a pulse. However,
when triggering of pulses can be automated, a function accepting as only
argument the number of pulses can be passed as argument to parameter
`f.trigger.pulses`. **It is important to be aware that no multitasking
support is provided, so for automation, the pulse trigger function
should return to the caller before the pulses are triggered and the
train of pulses start with enough delay for the integration to be
started before them.**

### Spectral transmittance, absorptance and reflectance

**Several features available in `acq_irrad_interactive()` have not been
yet implemented in function `acq_fraction_interactive`.**

A single function can be used for the acquisition of data where the
quantities of interest are expressed relative to a reference. This is
the case for transmittance, reflectance and absorptance. The same
parameters as discussed for `acq_irrad_interactive()` iare available in
function `acq_fraction_interactive()`. We will here describe three
additional parameters only present in this function. The first one is
`ref.value` which needs to be supplied if the reference target is
imperfect, such as a white reference patch which reflects less than
100%. The default is `ref.value = 1`, but this parameter not only
accepts numeric values as argument but also transmittance or reflectance
spectra for the target used as reference. Consequently if the
reflectance of an "imperfect" reference is know, such reference can be
used. (e.g. In the case of measuring the reflectance of very dark
objects, it might be convenient to use a grey reference instead of a
white reference). For a reference with 97% reflectance across the whole
range of wavelengths of interest one would use.

```{r, eval=sr.online}
acq_fraction_interactive(ref.value = 0.97,
                         correction.method = ooacquire::MAYP11278_ylianttila.mthd,
                         descriptors = ooacquire::MAYP11278_descriptors)
```

Parameters `qty.out` and `type` determine the class of object and the
quantity stored in the returned objects. If the optical set-up used is
for measuring specular spectral reflectance, we use the following code.
This information will also be used when plotting the data.

```{r, eval=sr.online}
acq_fraction_interactive(qty.out = "Rfr", type = "specular",
                         correction.method = ooacquire::MAYP11278_ylianttila.mthd,
                         descriptors = ooacquire::MAYP11278_descriptors)
```

Wen using *light sources with pulsed emission* we start a session by
calling function `acq_fraction_pulsed_interactive()` instead of
`acq_fraction_interactive` and we set manually the integration time as
when measuring spectral fluence. In other respects the procedure is
similar to that for continuous light sources except for the need to
either manually or automatically trigger the light pulses.

## Rolling your own

The functions for interactive acquisition of spectral data described
above, although useful in many cases, are meant also as examples. In
cases of routine measurements one could simplify menus by removing
entries that are not needed, or even converting the functions to
non-interactive, and possibly adding support for scheduled sequences of
spectral data acquisitions. This should be fairly simple, not requiring
advanced knowledge of the R language as we export the lower level
functions that can be used as building blocks for new variations. It
should be even possible to use these functions to develop a user
interface based of package 'shinny'.

## Using scripts

It is also possibly to combine lower level functions using R scripts
instead of through writing higher level R functions from them. Several
example scripts are included with the package. I recommend that you open
them, and save them to your own working directory before sourcing them
or editing them.

You will need first to locate where in your computer's file system the
'ooacquire' package is installed, and where the file you are interested
in is, and to copy it to your own workspace. Here we copy the script
called `"irrad-acq-interac.R"`.

```{r, eval=FALSE}
filepath <- system.file("example-scripts", "irrad-acq-interac.R", package="ooacquire")
file.copy(from = filepath, to = ".")
```

Here we copy all the example scripts.

```{r, eval=FALSE}
folderpath <- system.file("example-scripts", package="ooacquire")
list.files(path = folderpath, pattern = ".*[.]R")
file.copy(from = list.files(path = folderpath, pattern = ".*[.]R", full.names = TRUE), to = ".")
```
