---
title: "Default Arguments for Data Acquisition"
subtitle: "`ooacquire` `r packageVersion('ooacquire')`"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
---

# Introduction

The default behaviour of functions `acq_irrad_interactive()` and `acq_fraction_interactive()` can be modified by overriding the default argument values when calling them and by setting R options. This vignette describes what is possible and how to achieve it.

Using R options in ways that alter the behaviour of functions is in general discouraged because it compromises reproducibility. However, in functions for interactive use, such as those for the acquisition of spectral data, using R options to adjust the user interface or the style of live plots is usually safe as the saved data are not affected. Some options supported by 'ooacquire', do alter how data are expressed, e.g. units and/or basis of expression, but the returned objects always contain enough information stored as metadata attributes to transparently re-express the data as if it had been acquired using any of the available settings for the options.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ooacquire)
```

# Verbosity of messages and warnings

R option `verbose`, used by R itself, is respected by `ooacquire` and other R packages in the R for Photobiology Suite. In contrast, R option `photobiology.verbose` affects only the packages in the R for Photobiology suite. The default value for `verbose` is `FALSE` and that for `photobiology.verbose` the value of R option `verbose` if set, and otherwise `FALSE`.

A logical value passed as argument to parameter `verbose` in `acq_irrad_interactive()`, `acq_fraction_interactive()` and in other functions overrides the value of the corresponding R option for the messages and warnings issued by each function but not others. In most cases when `verbose` is `TRUE` messages and warnings are more detailed, and additional ones may be issued to alert about _possible_ problems or to _highlight_ actions taken automatically.

_Verbose output is useful when writing scripts, when problems are encountered, or when trying to better understand what a function is doing._

# Sanity checks on spectra

Functions from the R for Photobiology suite of packages, including `ooacquire`, check the validity of the spectral data when objects are created and manipulated. This adds overhead to computations but makes it easier to detect bad or corrupted data early. The checks are implemented in package 'photobiology' and how stringent they are when checking for out-of-range values is controlled by R option `photobiology.strict.range`, set to `TRUE`, `FALSE`, or `NA`, with `NA` fully disabling the range checks, and `FALSE` for less strict checks. The default value is `FALSE`.

_Which level to use depends on the properties of the data being acquired. Using `NA` to disable checks is not recommended as it may lead to difficult to diagnose errors in subsequent computations. Use of `TRUE` can help track the origin of unusual data._

# Units and basis of expression of radiation

The default unit for irradiance is $W m^{-2}$, as for the rest of the R for Photobiology suite of packages. This default can be changed, using functions from package 'photobiology', into $\mathrm{mole}\,s^{-1}\,m^{-2}$ by calling function `photon_as_default()` and the change undone by calling `energy_as_default()` or `unset_radiation_unit_default()`. Concurrently these same calls, change the default basis of expression for fluence from $\mathrm{J\,m^{-2}}$ per exposure event to $\mathrm{mole}\,m^{-2}$ per event and back. The switch between energy and photon units is possible also through a menu in the user interface, while plots as displayed.

_This option affects both plots displayed and how the spectral data is expressed in the objects returned and saved to disk. As interconversion is possible and done transparently, the choice is mostly dependent the desired expression in plots during acquisition._

# Measured quantity

In a call to `acq_irrad_interactive()` the argument passed to parameter `qty.out` selects between `"irradiance"`, `"fluence"`, `"cps"` (counts per second) or `"raw"` (raw detector counts) for the expression of the measured spectra, and also adjusts the user interface to match.

Irradiance is expressed per time unit, it is a flux rate, while fluence is expressed per measured event, i.e., integrated through time. Counts-per-second are raw-counts from the spectrometer analogue-to-digital converted divided by the integration time.

_As the choice in this case is based on the type of measurement and affects the quantity returned in a fundamental way, this choice cannot be set with an R option. An explicit argument in the function call is always needed._

# Quantities used for optical properties

In this case the default is to use reflectance, absorptance and transmittance as fractions adding up to 1. Because they add up two 1, specifying two of them determines the value of the third one. When comparing objects with vastly different properties or in chemistry when estimating solute concentrations is usual to use absorbance, which is based either on a $log_{10}$ or $log_e$ transformation. In the R for Photobiology Suite $log_{10}$ is always used for absorbance. The default can be changed to absorbance by calling `A_as_default()`, to absorptance by calling `Afr_as_default()` and to transmittance by calling `Tfr_as_default()` or `unset_filter_qty_default()`.

In a call to `acq_fraction_interactive()` the argument passed to parameter `qty.out` selects between `"Tfr"` (transmittance), `"Rfr"` (reflectance) or `"raw"` (raw detector counts) for the expression of the measured spectra, and also adjusts the user interface to match.

Which quantity is being measured is determined by the optical set-up used for measurement, and should always be set based on this.

_As the choice in this case is based on the type of measurement and affects the quantity returned in a fundamental way, this choice cannot be set with an R option. An explicit argument in the function call is always needed._

**Conversions are only possible when at least two of the quantities are known, so in some cases they are not available.**

# Time units used for fluxes

For irradiance the default unit for time is seconds, and is computed based on the integration time used for the measurement. For fluence, transmittance and reflectance, which are not fluxes, but a flow or fractions, time units are not relevant.

_When measuring spectral data there is rarely need to use a different time unit. Longer time units such as day, are used for daily totals calculated by integration rather than directly measured._

# Plotting defaults

To avoid bloating the interface of the spectral-data-acquisition functions and to avoid a tight dependency on the plotting functions, the control of plotting style, annotations and wavelength range are all controlled through R options. Some of these options can also be set through menus in the user interface. None can be set by passing arguments in the function call.

Plotting is done by calling `autoplot()` method specializations from package 'ggspectra'. Several default arguments for these methods can be modified by setting R options, in most cases both with R's function `option()` and with simpler wrapper functions from package 'ggspectra'.

Please see the documentation of functions `ggspectra::set_annotations_default()` (option `"photobiology.plot.bands"`), and `ggspectra::set_w.band_default()` (option `"photobiology.plot.annotations"`) and of `ggspectra::autoplot()` methods for details. It is also possible to interactively chose some common sets of wavebands through a menu.

"photobiology.plot.bands"
"ggspectra.wlrange"
"photobiology.radiation.unit"
"ggspectra.pc.out"
"photobiology.plot.annotations"
"photobiology.filter.qty"
"photobiology.reflector.qty"
"ggspectra.axis.symbols"
"photobiology.fancy.chars"
"photobiology.verbose"
"photobiology.use.hinges"
"photobiology.waveband.trim"
"photobiology.Tfr.zero"
"photobiology.strict.range"
"photobiology.check.spct"
"photobiology.use.cached.multipliers"
"photobiology.brief.trunc.names"

"digits"
"width"

"verbose_as_default" ???! in spct.normalize()

#Ignored
"photobiology.math"
"photobiology.object.qty"

Options that can be set directly with `options()` are `"ggspectra.pc.out"` (fractions as percentages) and `"ggspectra.wlrange"` (range of wavelengths to plot). The default range is that for which an irradiance calibration is available in the case of irradiance and fluence, and otherwise the full wavelength range of the spectrometer. It is also possible to interactively zoom-in by entering numeric values for a narrower range than that of the data.

Options `"photobiology.fancy.chars"` defines the mapping of SI prefixes to characters (such as k for kilo), `"ggspectra.axis.symbols"` can be used to disable/enable the use of symbols for physical quantities in axis labels.

 The default theme can be set with function `set_theme()` from package 'ggplot2', keeping in mind that the theme set as default can be modified with a call to `theme()`. One is not limited to the themes defined in package 'ggplot2', but can also use full themes from other packages. Themes from package 'ggdark' are specially useful when doing measurements in a dark room. A dark theme can be used on screen while the matching normal theme is used when saving the PDF files. The dark screen-only theme, can be enabled interactively through the menu.
 
In some cases if the R option is not set (`NULL`) these functions override the defaults used by setting the option. This annotation is slow to render, and is by default not included in plot to improve performance.
